---
title: "Logical_Operator"
output: html_document
---

How many patients are younger than 40
``` {r}
sum(age < 40)

```
How many patients are older than 25, and have an FEV1 between 40 and 70 (inclusive) 
``` {r}
age_logical <- age > 25
fev_logical <- (fev1 >= 40) & (fev1 <= 70)
sum(age_logical & fev_logical) 

```
How many patients have an even age (divisible by 2) and an odd fev1? 
``` {r}
sum((fev1 %% 2 == 1) & (age %% 2 == 0))

```
``` {r}
reduced_lungfunction <- fev1 < 70 

print(reduced_lungfunction)

names(fev1)[reduced_lungfunction
            ]

```

If statements
``` {r}


```

Else Statments 
``` {r}



```

Else-if Statements
``` {r}


```

Else-if Practice 
"Healthy" if FEV1 is greater than 80 
"Moderate impairment" if FEV1 between 60 and 80 
"Poor" if FEV1 is between 40 and 60 
"Very poor" if FEV1 is less than 40
``` {r}
patient_2_fev1 <- fev1["Patient_02"]

if (patient_2_fev1 < 80) {
  print("Healthy")
} else {
  print("Unhealthy")
} else if (patient_2_fev1 > 60) {
  print("Moderarte Impairment")
} else if (patient_2_fev1 > 40) {
  print("Poor")
} else 
  print("Very poor")
}


```
``` {r}
for (patient in fev1) {
  if (patient > 80) {
    print("Healthy")
  } else if (patient > 40) {
    print("Poor")
  } else {
    print("Very Poor")
  }
}

  

```
Paste 
Now modify the loop to print something like:Patient_01: HealthyPatient_02: Moderate
``` {r}


patient_2_fev1 <- fev1["Patient_01"]
patient_2_fev1 <- fev1["Patient_02"]
  
for (patient_id in names (fev1)) {

  if (fev1[patient_id] > 80) {
    print(paste0(patient_id, " : Healthy"))
  } else if (fev1[patient_id] > 60) {
    print(paste0(patient_id, " : Moderate impairment"))
  } else if (fev1[patient_id] > 40) {
    print(paste0(patient_id, " : Poor"))
  } else {
    print(paste0(patient_id," : Very poor"))
  }
}




```
Write a for loop that prints every age
``` {r}

for (patient_age in age) {
  print(patient_age)
}

```
Modify the for loop to print the patient id and age 
``` {r}
for (patient_age in names(age)) {
  print(paste(patient_age,age[patient_age]))
}

```
``` {r}
total_age <- 0

for (id in names(age)) {
  total_age <- total_age + age[id]
  
  print(paste(id, " FEV:", fev1[id], " Age:", age[id]))
  
  if (total_age >= 800) {
    break
  }
}


```



``` {r}
over_30_vec <- c()

for 

```