---
title: "Functions"
output: html_document
---
```{r}
fev1_bcch <- c(
  Patient_01 = 85,
  Patient_02 = 62,
  Patient_03 = 45,
  Patient_04 = 91,
  Patient_05 = 38,
  Patient_06 = 70,
  Patient_07 = 55,
  Patient_08 = 76,
  Patient_09 = 49,
  Patient_10 = 88
)
age_bcch <- c(
  Patient_01 = 23,
  Patient_02 = 62,
  Patient_03 = 45,
  Patient_04 = 91,
  Patient_05 = 38,
  Patient_06 = 70,
  Patient_07 = 55,
  Patient_08 = 76,
  Patient_09 = 49,
  Patient_10 = 88
)

fev1_sph <- c( 
Patient_01 = 31, 
Patient_02 = 54, 
Patient_03 = 29, 
Patient_04 = 78, 
Patient_05 = 41, 
Patient_06 = 66,
Patient_07 = 35, 
Patient_08 = 82, 
Patient_09 = 57, 
Patient_10 = 46
)

age_sph <- c(
  Patient_01 = 72,
  Patient_02 = 58,
  Patient_03 = 91,
  Patient_04 = 44,
  Patient_05 = 67,
  Patient_06 = 83,
  Patient_07 = 39,
  Patient_08 = 76,
  Patient_09 = 52,
  Patient_10 = 61
)





```

Using the BCCH data, write a loop that creates a character vector called risk_bcch containing a risk classification for each patient.
Assign patients as:
"High risk" if: 
Age is greater than 50 and 
FEV1 is less than 70 
"Moderate risk" if: 
Age is greater than 30 or 
FEV1 is less than 70 
"Low risk" otherwise 
The final vector should:
Have the same length as the input data 
Use patient IDs as names


```{r}
#' Classifying risk for patients based on age and fev1
#'
#' @param fev1 fev1 is a numeric vector
#' @param age age is a numeric vector
#' 
#' 
#' 
#' @return none 
#'
```


```{r}
#'

for (id in names(age_bcch)) {
  if (age_bcch[id] > 50 & fev1_bcch[id] < 70) { 
    print(" High Risk")
} else if (age_bcch[id] > 30 & fev1_bcch[id] < 70) {
  print("Moderate Risk")
} else { 
  print(" Low Risk ")
  }

}

classify_risk <- function(fev1, age) {
  for (id in names(age)) {
  if (age[id] > 50 & fev1[id] < 70) { 
    print(" High Risk")
} else if (age[id] > 30 & fev1[id] < 70) {
  print("Moderate Risk")
} else { 
  print(" Low Risk ")
  }
  
  }
} 


classify_risk(fev1_bcch, age_bcch)

print("SPH")
classify_risk(fev1_sph, age_sph)


```

```{r}
install.packages("palmerpenguins")

```

```{r}
library(palmerpenguins) 

```


```{r}
data("penguins")

```

```{r}

write.csv( 
penguins, 
"penguins.csv", 
row.names = FALSE 
)

```

```{r}
penguins <- read.csv("penguins.csv")

```