---
title: "Practice"
output: html_document
---

```{r}
amrit <- 80 
sam <- 90

average <- mean(c(amrit, sam))
print(amrit) 
print(sam)
print(mean(c(amrit,sam)))
print <- (average)

```
A teacher recorded the marks of three students.
Emma scored 72
Jack scored 85
Olivia scored 91
Your tasks
Create variables called emma, jack, and olivia.
Create a new variable called scores that stores all three marks.
Create a variable called average that stores the average score.
Print:
emma
scores
average
```{r}
emma <- 72
jack <- 85
olivia <- 91

scores <- c(emma,jack,olivia)
average <- mean(scores)
print(emma)
print(scores)
print(average)



```

``` {r}

name <- "maria"
university <- "university of glasgow" 
degree <- "biochemistry" 

student_info <- c(name,university,degree)
print(name)
print(student_info)

message <- "Welcome to R"
print(message)

```
```{r}

animals <- c("lion","zebra","elephant")
print(animals[2])

```
A coffee shop sold the following number of drinks today:
Monday: 35
Tuesday: 42
Wednesday: 38
Thursday: 47
Your tasks
1. Create four variables:
monday
tuesday
wednesday
thursday
```{r}
monday <- 35
tuesday <- 42
wednesday <- 38
thursday <- 47

sales <- c(monday,tuesday,wednesday,thursday)

average_sales <- mean(sales)
print(sales)
print(average_sales)
print(sales[3])

sum(sales)
print(sum(sales))
```
This is something you'll use all the time.
Question 4
A shop sold:
Apples = 12
Bananas = 18
Oranges = 15
Your tasks
Create the three variables.
Create a vector called fruit_sales.
Print the vector.
The shop sold 5 more apples. Update the variable apples.
Recreate the fruit_sales vector so it uses the updated value.
Print the updated vector.
Print the new total sales.
```{r}
apples <- 12
bananas <- 18
oranges <- 15

apples <- apples + 5

fruit_sales <- c(apples,bananas,oranges)


print(fruit_sales)




```
A teacher recorded the following test scores:
Alice = 78
Ben = 65
Chloe = 91
David = 84
Emma = 73
Your tasks
Create a variable for each student.
Create a vector called:
scores
Print the vector.
Calculate and store the average in a variable called:
average_score
Print the average.
Print only Chloe's score.
Print the highest score (without looking at the numbers yourself—see if you can guess the function).
Print the lowest score (again, think about what the function might be called).
```{r}

alice <- 78
ben <- 65
chloe <-  91
david <- 84
emma <- 73

scores <- c(alice,ben,chloe,david,emma)
print(scores)
average_scores <- mean(scores)
print(average_scores)
print(scores[3])
highest_score <- max(scores)
print(highest_score)
lowest_score <- min(scores)
print(lowest_score)
```
We're going to start using comparison operators.
Question 6
Using the same scores vector:
Print all the scores that are greater than 80.
Print all the scores that are less than 75.
Create a variable called:
passed
that stores whether each student scored 70 or above.
Print passed.
💡 Don't worry if you don't know how to do it. Think back to when we used indexing:
scores[3]
Can you think of a way to put a condition inside the square brackets?
Have a go without looking anything up. Even if it's wrong, that's how you'll learn.

``` {r}
print(scores[scores > 80])
print(scores[scores < 75])
passed <- scores >= 70


```
Question 7 – Patient Heart Rates
A hospital records the following heart rates (beats per minute):
Patient 1 = 62
Patient 2 = 88
Patient 3 = 75
Patient 4 = 102
Patient 5 = 69
Patient 6 = 81
Your tasks:
Create variables for each patient.
Create a vector called heart_rates.
Print the vector.
Print the average heart rate.
Print only the heart rates above 80 bpm.
Create a variable called high_heart_rate that stores whether each patient has a heart rate above 100 bpm.
Print high_heart_rate.
Print only the heart rates that are above 100 bpm.
``` {r}
patient_1 <- 62
patient_2 <- 88
patient_3 <- 75
patient_4 <- 102
patient_5 <- 69
patient_6 <- 81

heart_rates <- c(patient_1, patient_2, patient_3, patient_4, patient_5, patient_6)
print(heart_rates)
average_rate <- mean(heart_rates)
print(average_rate)
print(heart_rates[heart_rates > 80])
high_heart_rate <- heart_rates > 100
print(high_heart_rate)
print(heart_rates[heart_rates > 100])




```
Using your heart_rates vector:
Print the patient(s) whose heart rate is exactly 75.
Print the patient(s) whose heart rate is not equal to 75.
Create a variable called:
normal_rate
that stores whether each patient has a heart rate between 60 and 100 inclusive.
⚠️ This last part is the only new challenge. Think about it before you code.
Hint: You'll need two comparisons joined together.
Give it your best shot. Don't worry if the last one isn't quite right—we'll work through it together.
``` {r}
print(heart_rates[heart_rates = 75])
print(heart_rates[heart_rates != 75])
normal_rate <- heart_rates >= 60 & heart_rates <= 100
print(normal_rate)

```
A researcher records the ages of six participants:
ages <- c(18, 22, 17, 25, 19, 16)
They define an adult as someone aged 18 or older.
Without creating any new vectors manually, answer the following:
Print all participants who are adults.
Print all participants who are minors.
Create a variable called adult that stores whether each participant is an adult.
Print adult.
How many adults are there?
```{r}
ages <- c(18, 22, 17, 25, 19, 16)
print(ages[ages >= 18])
print(ages[ages <18 ])
adult <- ages >= 18
print(adult)
sum(adult)


```
A researcher records a participant's blood glucose level.
glucose <- 6.8
Write an if/else statement that:
prints "Normal" if the glucose level is less than 7
otherwise prints "High"

``` {r}
glucose <-  6.8
if (glucose < 7 ) {
  print("Normal")
} else {
  print("High")
}

```
Question 2 — if / else if / else
A student receives a mark.
mark <- 68
Write an if statement that prints:
"Distinction" if the mark is 70 or above
"Pass" if the mark is 50–69
"Fail" otherwise
```{r}
mark <- 68
if (mark >= 70) {
  print("Distinction") 
} else if (mark >= 50) { 
  print("Pass") 
} else { 
  print("Fail")
}
 



```
Question
A hospital records the heart rate of several patients:
heart_rates <- c(62, 88, 75, 102, 69, 81)
Using a for loop and an if/else statement, print:
"Normal" if the heart rate is 100 or below
"High" if the heart rate is above 100
The expected output is:
Normal
Normal
Normal
High
Normal
Normal
One hint
A for loop has this general structure:
for (item in vector) {

}
I want you to have a go at writing it from memory. Don't worry if you get the loop syntax slightly wrong—that's exactly what we're practising.
``` {r}
heart_rates <- c(62, 88, 75, 102, 69, 81)

for (heart in heart_rates)
  if (heart <= 100) {
    print("Normal")
  } else if (heart > 100) {
    print("High")
    
  }

```
Question 1 — Blood Pressure Classification
A doctor records the systolic blood pressure of six patients.
blood_pressure <- c(118, 145, 132, 155, 121, 168)
Using a for loop and an if/else if/else statement, print:
"Normal" if the blood pressure is less than 120
"Elevated" if the blood pressure is 120 to 139
"High" if the blood pressure is 140 or above
Expected output
Normal
High
Elevated
High
Elevated
High
``` {r}
blood_pressure <- c(118, 145, 132, 155, 121, 168)
for (bp in blood_pressure) {
  if (bp < 120) {
    print("Normal")
    
  } else if (bp <= 139) {
    print("Elevated")
    
  } else { 
     print("High") 
  }

}
  
    
blood_pressure <- c(118, 145, 132, 155, 121, 168)

for (bp in blood_pressure) {

  if (bp < 120) {
    print("Normal")

  } else if (bp <= 139) {
    print("Elevated")

  } else {
    print("High")
  }

}

```
You are given the following named vector:
temperatures <- c(
  London = 18,
  Glasgow = 15,
  Madrid = 31,
  Paris = 24,
  Rome = 33
)
Your task
Write a for loop that:
Loops through the city names.
Prints:
"Warm" if the temperature is 25°C or above.
"Cool" otherwise.
Prints the city name as well.
Your output should look like:
```{r}
temperatures <- c(London = 18, Glasgow = 15, Madrid = 31, Paris = 24,Rome = 33) 
  
for (city in names(temperatures)) { 
  if (temperatures[city] >= 25) {
    print(paste0(city , " : Warm" ))
  } else {
    print(paste0(city, " : Cool"))
  } 

}

  
  
  


```

Practice Question
heart_rate <- c(
  Alice = 72,
  Ben = 95,
  Chloe = 68,
  Daniel = 104,
  Emma = 81
)
Your task
Write a for loop that:
Loops through the patient names.
If the heart rate is 90 or above, print:
Ben : High
Otherwise print:
Alice : Normal
using paste0().
```{r}
heart_rate <- c(Alice = 72, Ben = 95, Chloe = 68, Daniel = 104, Emma = 81) 
                
for (patient in names(heart_rate)) {
  if (patient >= 90) {
    print(paste0(patient, " : High"))
  } else {
    print(paste0(patient, " : Normal"))
  }
    
  
  
}
  


```
You are given:
scores <- c(
  Alice = 65,
  Ben = 82,
  Chloe = 91,
  Daniel = 74,
  Emma = 56
)
Your task is to write a loop that:
Loops through each student's name.
Prints:
the student's name
their score
Keeps a running total of all the scores.
Stops the loop as soon as the running total is greater than 220.
```{r}
scores <- c(Alice = 65,Ben = 82,Chloe = 91,Daniel = 74,
Emma = 56)

total_score <- 0
for (student in names(scores)) {
  print(paste0(student, " : ", scores[student]))
  total_score <- total_score + scores[student] 
  if (total_score > 220)
    break
  
  
}


```


Your task
Before the loop, create a variable called:
total_steps <- 0
Then write a for loop that:
Loops through each person's name.
Prints the person's name and step count, for example:
Alice : 3500
Adds each person's steps to total_steps.
If the running total becomes greater than 15,000, stop the loop using break.
``` {r}
steps <- c(
  Alice = 3500,
  Ben = 8200,
  Chloe = 6100,
  Daniel = 4700,
  Emma = 9300
)
total_steps <- 0

for(person in names(steps)) {
  print(paste0(person, " : ", steps[person]))
  total_steps <- total_steps + steps[person]
  if (total_steps > 15000)
    break
}



```
A hospital has recorded the oxygen saturation (%) of five patients.
oxygen <- c(
  Alice = 98,
  Ben = 91,
  Chloe = 96,
  Daniel = 88,
  Emma = 94
)
Your task
Before the loop
Create two variables:
low_count <- 0
total_oxygen <- 0
Then write a for loop that:
Loops through each patient's name.
Adds each patient's oxygen level to total_oxygen.
If the oxygen level is below 92, print:
Ben : 91 (Low)
and increase low_count by 1.
Otherwise print:
Alice : 98 (Normal)
If low_count reaches 2, stop the loop immediately using break.
After the loop
Print:
Total oxygen = ...
using paste0().
```{r}
oxygen <- c(
  Alice = 98,
  Ben = 91,
  Chloe = 96,
  Daniel = 88,
  Emma = 94
)

low_count <- 0
total_oxygen <- 0

for (patient in names(oxygen)) {
  total_oxygen <- total_oxygen + oxygen[patient]
  if (oxygen[patient] < 92) {
    print(paste0(patient, " : " , oxygen[patient], " Low "))
    low_count <- low_count + 1
  } else {
    print(paste0(patient, " : ", oxygen[patient], " Normal ")) } 
  if (low_count == 2) {
  break }
  
} 
    
  
  

print(paste0("Total Oxygen = ", total_oxygen ))


```
Practice Assignment: 
You are given the following named vector:
scores <- c(
  Alice = 82,
  Ben = 47,
  Chloe = 91,
  Daniel = 65,
  Emma = 38,
  Frank = 74
)
Question 1 (2 marks)
Create a variable called average_score that stores the mean score.
```{r}
scores <- c(
  Alice = 82,
  Ben = 47,
  Chloe = 91,
  Daniel = 65,
  Emma = 38,
  Frank = 74
)
average_scores <- mean(scores)
print(average_scores)

```
How many students scored 70 or above?
Store the answer in a variable called:
```{r}
high_scores <- scores >= 70 
sum(high_scores)
print(high_scores)
print(sum(high_scores))

```
Create a logical vector called
passed
where a pass is 50 or above.
Print the logical vector.
```{r}
passed <- scores >= 50
print(passed)

```
Question 4 (5 marks)
Using a for loop, print every student's score like this:
Alice : 82
Ben : 47
```{r}
for (student in names(scores))
  print(paste0(student, " : ", scores[student]))

```
Question 5 (5 marks)
Modify your loop.
If a student passed, print:
Alice : Pass
Otherwise print:
Ben : Fail
```{r}
for (student in names(scores)) {
  if (scores[student] >= 50)
    print(paste0(student, " : Pass " )) 
 else {
   print(paste0(student, " : Fail " ))
 }
}
```
Question 6 (6 marks)
Before the loop create
pass_count <- 0
Every time a student passes:
increase pass_count by 1.
After the loop print
Total passes = 4
```{r}
pass_count <- 0
for (student in names(scores)) {
  if (scores[student] >= 50) {
    print(paste0(student, " : Pass " )) 
  pass_count <- (pass_count + 1) 
  } else {
   print(paste0(student, " : Fail " ))
  }
  
}

print(paste0( " Total passes = ", pass_count))
```
Question 7 (8 marks)
Before the loop create
total_score <- 0
Inside the loop:
add each student's score to total_score
print whether they passed or failed
stop the loop once total_score becomes greater than 250
(Hint: you'll need break.)
After the loop print
Running total = ...
```{r}
total_score <- 0
for (student in names(scores)) {
  if (scores[student] >= 50) {
    print(paste0(student, " : Pass " )) 
  } else {
   print(paste0(student, " : Fail " ))
  }
  total_score <- total_score + scores[student]
  if (total_score > 250)
    break 
  
}


print(paste0(" Running total = ", total_score))


```
Question 8 (9 marks)
Create a new named vector:
heart_rate <- c(
  Alice = 72,
  Ben = 115,
  Chloe = 68,
  Daniel = 102,
  Emma = 77
)
Write a program that:
creates
high_count <- 0
and
total_hr <- 0
Then loops through every patient.
For each patient:
add their heart rate to total_hr
if heart rate is 100 or above
print
Ben : High
increase high_count
otherwise print
Alice : Normal
If high_count reaches 2, stop the loop.
After the loop print:
High heart rates = ...
Total heart rate = ...
```{r}
heart_rate <- c(
  Alice = 72,
  Ben = 115,
  Chloe = 68,
  Daniel = 102,
  Emma = 77
)
high_count <- 0 
total_hr <- 0

for (patient in names(heart_rate)) {
  total_hr <- total_hr + heart_rate[patient] 
  if (heart_rate[patient] >= 100) {
    print(paste0(patient, ": High"))
    high_count <- high_count + 1 
  } else {
    print(paste0(patient, " : Normal ")) }
  if (high_count == 2)
    break

}

print(paste0("High heart rates = ", high_count))
print(paste0("Total heart rates = ", total_hr))

```
# Question 9

# You are given the following named vector:

temperature <- c(
  Alice = 36.8,
  Ben = 39.2,
  Chloe = 37.1,
  Daniel = 38.7,
  Emma = 36.5,
  Frank = 39.5
)

# Write a program that:

# 1. Creates:
# fever_count <- 0
# total_temperature <- 0

# 2. Loops through every patient.

# 3. For each patient:
#    - Add their temperature to total_temperature.
#    - If temperature >= 38.0:
#         Print: "Ben : Fever"
#         Increase fever_count by 1.
#      Otherwise:
#         Print: "Alice : Normal"

# 4. Stop the loop when fever_count reaches 2.

# 5. After the loop, print:
#    Number of fevers = ...
#    Total temperature = ...
```{r}
temperature <- c(
  Alice = 36.8,
  Ben = 39.2,
  Chloe = 37.1,
  Daniel = 38.7,
  Emma = 36.5,
  Frank = 39.5
)

fever_count <- 0
total_temperature <- 0

for (patient in names(temperature)) {
  total_temperature <- total_temperature + temperature[patient] 
  if (temperature[patient] >= 38.0 ) { 
    print(paste0(patient, " : Fever "))
    fever_count <- fever_count + 1
    } else { 
      print(paste0(patient, " : Normal ")) }
    if (fever_count == 2) 
      break 

  
  
} 
print(paste0(" Number of Fevers = ", fever_count))
print(paste0(" Total Temperature = ", total_temperature))

sum(fever_count)
sum(total_temperature)





```

